### 2015-01-23 mcmc.tex
notetoself
typesetting code; advice, notes to self, notation

% This file is part of the MCMC project,
%   which is part of the Data Analysis Recipes project
% Copyright 2013, 2014, 2015 David W. Hogg (NYU) and Dan Foreman-Mackey (NYU).

@@

% style notes
% -----------
% - These documents are ``Chapters'', no?
% - [LaTeX] newline after every full stop for proper git diffing
% - [LaTeX] eqnarray not equation
% - pdf not PDF
% - is it ``a MCMC sampling'' or ``an MCMC sampling''?
% - \note{} right after word if you are endnoting a phrase,
%   but after the full stop if endnoting the sentence or paragraph.
% - do we work in $x$ or $\pars$?  I am confused...
% - should we use the word ``walker'' anywhere, everywhere, or never?

@@

Heuristically, a sampling $\setof{x_k}_{k=1}^K$ from some pdf $p(x)$
is a set of $K$ values $x_k$ that are independent draws from the pdf.
Heuristically, if an enormous (large $K$) sampling could be displayed
in a finely binned $x$-space histogram, the histogram would look---up
to a total normalization---just like the original function $p(x)$.

Not being mathematicians, we don't know the precise definition of a
pdf\note{Seriously.  We don't.  It has something to do with measure theory.}, but for our purposes here a pdf is
any single-valued (scalar) function that is non-negative everywhere
(the entire domain of $x$), and obeys a normalization condition
\begin{eqnarray}
0 &\leq& p(x) \quad \mbox{for all $x$}
\\ \label{eq:normp}
1 &=& \int p(x)\,\dd x
\quad ,
\end{eqnarray}
where, implicitly, the integral is over the full domain of $x$.
Importantly, although $p(x)$ is single-valued, $x$ can be a scalar or
a multi-element vector or list or blob; it can be arbitrarily large
and complicated.
In data-analysis contexts, $x$ will often be the full blob of free
parameters in the model.
Implicitly, the integral in \equationname~(\ref{eq:normp}) is high-dimensional; it
has as many dimensions as there are elements or entries or components of $x$.
Also, if there are elements or entries or components of $x$ that are discrete
(that is, take on only integer values or equivalent),
then along those dimensions the integral becomes a discrete sum.
This latter is a detail to which we return below.

Given this pdf $p(x)$, we can define \emph{expectation values} $E[x]$
for $x$ or for any quantity that can be expressed as a function $q(x)$
of $x$:
\begin{eqnarray}
E[x] &\equiv& \int x\,p(x)\,\dd x
\\
E[q] &\equiv& \int q(x)\,p(x)\,\dd x
\quad ,
\end{eqnarray}
where again the integrals are implicitly definite integrals over the
entire domain
of $x$ (all the parts of $x$ space in which $p(x)$ is finite) and the
integrals are multi-dimensional if $x$ is multi-dimensional.
These expectation values are the mean values of $x$ and $q(x)$ under
the pdf.  A good sampling---really the definition of a good
sampling---makes the sampling approximation to these integrals
accurate.
With a good sampling $\setof{x_k}_{k=1}^K$ the integrals get replaced
with sums over samples $x_k$:
\begin{eqnarray}
E[x] &\approx& \frac{1}{K}\,\sum_{k=1}^K x_k
\\
E[q] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(x_k)
\quad .
\end{eqnarray}
That is, a sampling is good when any expectation value of interest is
accurately computed via the sampling approximation.
The word ``accurately'' here translates into some kinds of theorems
about limiting behavior; the general idea is that the sampling
approximation becomes exact as $K$ goes to infinity.
The size $K$ of the sampling you need \emph{in practice} will depend
on the expectations you want to compute, and the accuracies you need.

As we noted above, in the context of MCMC, we are often sampling or
using some \emph{badly normalized} function $f(x)$.
This function is just the pdf $p(x)$ multiplied by some unknown and
hard-to-compute scalar.
In this case, for our purposes, the conditions on $f(x)$ are that it
be non-negative everywhere and have \emph{finite} integral $Z$
\begin{eqnarray}
0 &\leq& f(x) \quad \mbox{for all $x$}
\\
Z &=& \int f(x)\,\dd x
\quad .
\end{eqnarray}

When the sampling $\setof{x_k}_{k=1}^K$ is of one of these badly
normalized functions $f(x)$---as it usually will be---the
sampling-approximation expectation values are the expectation values
under the properly-normalized corresponding pdf, even though you might
\emph{never learn that normalization}.
When we run MCMC sampling on $f(x)$, a function which differs from a
pdf $p(x)$ by some unknown normalization constant, then the sampling
permits computation of the following kinds of quantities:
\begin{eqnarray}
E[q] &\equiv& \frac{\int q(x)\,f(x)\,\dd x}{\int f(x)\,\dd x}
\\
E[q] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(x_k)
\quad .
\end{eqnarray}
That is, the sampling can be constructed (as we will show below) from
$f(x)$ directly, and it permits you to compute expectation values
without ever requiring you to integrate either the numerator integral
or the denominator integral, both of which are generally
intractable.\note{As we will
  see, the intractability comes from various directions, but one is
  that the dimension of $x$ gets large in most realistic situations.
  Another is that the support of $p(x)$ tends to be, in normal
  inference situations, \emph{far smaller} than the full domain of
  $x$.
  That is, the pdf is at or very close to zero over all but some tiny
  and very hard-to-find part of the space.}

@@

% in the next paragraph I adopt $d$ to be the dimensionality of the space

For example, one magical property of a sampling (in a $d$-dimensional space)
  is that a \emph{histogram} (a $d$-dimensional histogram) of the samples
  (divided, if you like, by the number of samples in each bin and the bin width\note{something about units and normalization})
  looks very much like the pdf from which the samples were drawn.
This is a way to ``reconstruct'' the pdf from the sampling:
Make a histogram from the samples.
Even in this case, the sampling is being used to do \emph{integrals};
  the (possibly odd) idea is that the approximate or effective value of the pdf in each bin
  of the histogram is an average over the bin.
That average is obtained by performing an integral.

@@

Finally---and perhaps most importantly---a magical property of a sampling
  (in a $d$-dimensional space)
  is that if some of your $d$ dimensions are totally uninteresting
  (nuisance parameters, if you will)
  and some of your $d$ dimensions are of great interest,
  the sampling in the full $d$-space is trivially converted into a sampling in the subspace of interest:
You just drop from each $x_k$ vector (or blob or list) the dimensions of no interest!
That is, the projection of the sampling to the subspace of interest produces
  a sampling of the marginalized pdf, marginalizing (or projecting) out the nuisance parameters.%
  \note{This point about marginalization is, once again, a use of the sampling to perform an \emph{integral};
        the marginalized pdf is obtained from the full pdf by an integration.
        The sampling performs this integration automatically.}
That is extremely important for inference,
  where there are always parameters with very different levels of importance to the scientific conclusions.
This point will return again below, either explicitly or implicitly. % HOGG: does it?

@@

In what follows, when we discuss tuning of the proposal pdf, we will
  often cycle through the parameter blob components and propose changes
  in only one dimension or element or component at a time.
That is, at step $k+1$ you might only propose a one-dimensional move
  in the $i$th dimension, not in all $d$ dimensions of the parameter
  space.
That won't change anything significant; but if you are implementing
  this right now you might want to do that.
It will help us tune the proposal pdf (\sectionname~\ref{sec:tuning})
and diagnose problems (\sectionname~\ref{sec:troubleshooting}).

@@

One note to make here
  is that you must \emph{propose}
  in the same coordinates (parameterization or transformation of parameters)
  as that in which your priors are \emph{specified},
  or else multiply in a (possibly nasty) Jacobian.
That is, if your prior is ``flat in $x$''
  but you find it easier to run your sampler by proposing steps in $\ln x$,
  if you don't modify your acceptance fraction by the Jacobian,
  your \emph{real} prior won't be flat in $x$.
These think-os can get subtle;
  the best way to avoid them
  is to write your code and your likelihood function
  and your prior pdf function and your proposal distribution
  all in precisely the same parameterization.
We will return to this point below in our comments on functional testing.

@@

Most MCMC methods
  (though not all\note{We are promoters of an ensemble method that has a
  nearly-fixed proposal distribution and requires no tuning (CITE EMCEE).})
  make use of something like a ``proposal distribution''
  which determines what kinds of steps the walker can take
  as it random-walks through the parameter space.
The user generally has a lot of control over what this proposal distribution might be,
  and how to choose the parameters.
For example, in a $d$-dimensional parameter space, %do we consistently use d here?
  a zero-mean but anisotropic Gaussian (normal distribution)
  is often used to draw offsets for the walker to take from point to point.
Even within this choice (which is of one function among many possibilities),
  there are $d\,(d+1)/2$ parameters in the $d\times d$ symmetric, positive definite covariance matrix
  to set ``by hand''.
How to choose this function and set these parameters?

@@

A very common---%
  and very useful---%
  kind of proposal distribution is one that cycles through parameters,
  taking a random step in just one parameter at a time.
Aside from involving some persistence of state,
  this kind of proposal distribution can be valuable,
  in part because it reduces the (almost impossible) $d$-dimensional tuning problem
  to $d$ one-dimensional problems:
In this form of proposal distrubution,
  it is possible to track a separate acceptance fraction for every parameter.
Code can be built that uses the burn-in phase to tune all $d$ proposal variances
  such that each of them, individually,
  obtains acceptance at the same Goldilocks ratio.

@@

Another proxy for autocorrelation time useful for tuning is the
  Expected Squared Jump Distance.\note{(CITE GELMAN)}
This is the mean squared distance the walker moves, per step.
It is maximized when the acceptance ratio is reasonable
  and the step size is large;
  it is large when the exploration of the space is fast.
This proxy is easy to measure and use for tuning,
  and more directly related to autocorrelation time than the acceptance ratio.
We recommend using it for tuning, though we have never used it ourselves.
Like the acceptance fraction, it can also be used to tune
  in the case that the proposal distribution loops over parameters;
  once again, the user gets (in this case) $d$ one-dimensional tunings.

@@

As we mentioned above, scientists often like to report the ``best sample'';
  that is, the sample with the highest posterior pdf value.
This is not usually a good idea.
For one, MCMC is a sampling algorithm, not an optimization algorithm;
  there are no guarantees that it will find the optimum of the posterior pdf in reasonable time.%
  \note{The sampler samples the pdf fairly.
    The tallest peak in the posterior is not guaranteed to be---%
    and in general is not---%
    also the peak with the greatest posterior ``mass''.
    That is, the samples in a fair sampling will not necessarily come
    predominantly from the \emph{tallest} peak in the posterior pdf.
    They will come from the peak with the greatest total integral
    of the likelihood over the prior.
    These issues might sound ``academic'',
    but when the number of parameters gets large (greater than, say, 10),
    many normal intuitions one might have (if any) about what is reasonably likely in a pdf
    are regularly violated.}
For two, the closeness of the best sample to the posterior mode is a strong function of sample size,
  and with very bad scaling.%
  \note{If you have a $K$-point sampling in a $d$-dimensional parameter space,
    there is some best sample $\pars_k^{\best}$.  Now imagine that you want to make
    that best sample 10 times better---that is, you want to find a point $\pars^{\better}$
    that is ten times closer to the true optimum.  How many more samples do you need to take?
    In the limit that $K$ is large, on average, you have to take $10^d$ more samples!
    Most sensible optimizers are far, far better than this.}
For three, if you just want an optimum, use an optimizer!%
  \note{HOGG, something?}
That is, we \emph{strongly advise against} reporting,
  as ``the'' measurement or ``the result'',
  the parameter value or values for the best sample.
Giving some posterior samples is a good idea (see below).
If you feel drawn to give the best sample,
  instead give the parameter values found by optimizing the posterior PDF,
  starting at the best sample as an initialization
  (and sacrifice your probabilism\note{something}).
This optimal-posterior parameter value is called the
  ``maximum \foreign{a posteriori}'' or ``MAP'' value for the parameters;
  it is like a maximum-likelihood value but regularized by the prior.
It is an estimator with some good (and some bad) properties, but it is not the point of MCMC,
  and therefore outside the scope of this \documentname.

@@

MCMC is used to obtain samples $x_k$ from pdf $p(x)$,
  or a badly normalized all-positive function $f(x)$
  that is different from the pdf by an unknown factor $Z$.
In the context of data analysis,
  MCMC is usually being used to obtain samples
  that are \emph{parameter values} for a probabilistic model of some data.
That is, MCMC is sampling a pdf \emph{for the parameters}.
If it is returning parameter values, then it is not sampling a pdf for the data.
That is, MCMC can only sample a posterior (or prior) pdf for the parameters.
\emph{It cannot sample a likelihood.}%
\note{Well, technically, MCMC \emph{can} be used to sample a likelihood function,
  but the samples would be samples of possible \emph{data} not possible \emph{parameters}.
  The purist might say that even this is not sampling a likelihood function,
  because you should only call it a ``likelihood function'' in contexts in which you are
  treating the data as fixed (at the true data values) and the parameters as variable.
  In this context, the likelihood function is \emph{not} a pdf for the parameters,
  so it can't be sampled.}

@@

...Give a sense of the difficulties you get into if you fail here; like
discussion with Sven Kreiss on 2013-02-14...

...Explain how your code \emph{should look} with a prior function and a likelihood function...

@@

\paragraph{functional testing:}
In addition to a full set of unit tests for every part of your code
  (yeah \emph{right}\note{If you aren't unit testing, you are probably making some very big mistakes.}),
  we recommend building some end-to-end functional tests that permit you to perform MCMC
  runs with output that must meet certain expectations.
The simplest kind of functional test is to sample a known distribution,
  and check that the sampling has the moments you expect (to within tolerances).
For example, you can sample a $d$-dimensional Gaussian distribution with known non-trivial covariance,
  and check that the empirical covariance comes out as expected.
This is an easy test to fail, so a success builds confidence in any MCMC code.

In the most extreme form of this test,
  for a Gaussian in $d$ dimensions,
  build the covariance matrix input to the Gaussian distribution
  by taking the sum of the outer product of $d+1$ randomly generated vectors.
Then, before sampling, get the eigenvalues and eigenvectors of the matrix.
Set up the MCMC sampler with the Gaussian function with this covariance matrix
  as the probability function (or the log of it as the log-probability function).
Sample for many autocorrelation times,
  or until you are confident you are getting good samples.
Check that the empirical $d$-dimensional covariance of the samples
  has the same (or similar) eigenvectors and eigenvlues to the input matrix.
Also check that the individual one-dimensional samplings
  (along each axis or any projection in the $d$-space)
  look Gaussian and have nearly vanishing skew and kurtosis.
It takes some thinking and some calculating to see ``how well'' these things all ought to match
  expectations, but one thing that can be said with confidence is that agreement should improve
  as the sampling proceeds.
If it doesn't, it is likely that there is something broken about the sampler itself.

>>>>

% This file is part of the MCMC project,
%   which is (in turn) part of the Data Analysis Recipes project.
% Copyright 2013, 2014, 2015 David W. Hogg (NYU) and Dan Foreman-Mackey (NYU).

@@

% style notes
% -----------
% - These documents are ``Chapters'', no?
% - [LaTeX] newline after every full stop for proper git diffing
% - [LaTeX] eqnarray not equation
% - pdf not PDF
% - is it ``a MCMC sampling'' or ``an MCMC sampling''?
% - \note{} right after word if you are endnoting a phrase,
%   but after the full stop if endnoting the sentence or paragraph.
% - always $\pars$ never $x$; $K$ samples $\pars_k$ and $\pars$ space is $D$-dimensional
% - should we use the word ``walker'' anywhere, everywhere, or never?

@@

Heuristically, a sampling $\setof{\pars_k}_{k=1}^K$ from some pdf $p(\pars)$
is a set of $K$ values $\pars_k$ that are independent draws from the pdf.
Heuristically, if an enormous (large $K$) sampling could be displayed
in a finely binned $\pars$-space histogram, the histogram would look---up
to a total normalization---just like the original function $p(\pars)$.

Not being mathematicians, we don't know the precise definition of a
pdf\note{Seriously.  We don't.  It has something to do with measure theory.}, but for our purposes here a pdf is
any single-valued (scalar) function that is non-negative everywhere
(the entire domain of $\pars$), and obeys a normalization condition
\begin{eqnarray}
0 &\leq& p(\pars) \quad \mbox{for all $\pars$}
\\ \label{eq:normp}
1 &=& \int p(\pars)\,\dd\pars
\quad ,
\end{eqnarray}
where, implicitly, the integral is over the full domain of $\pars$.
Importantly, although $p(\pars)$ is single-valued, $\pars$ can be a scalar or
a multi-element vector or list or blob; it can be arbitrarily large
and complicated.
In data-analysis contexts, $\pars$ will often be the full blob of free
parameters in the model.
Implicitly, the integral in \equationname~(\ref{eq:normp}) is high-dimensional; it
has as many dimensions as there are elements or entries or components of $\pars$.
Also, if there are elements or entries or components of $\pars$ that are discrete
(that is, take on only integer values or equivalent),
then along those dimensions the integral becomes a discrete sum.
This latter is a detail to which we return below.

Given this pdf $p(\pars)$, we can define \emph{expectation values} $E[\pars]$
for $\pars$ or for any quantity that can be expressed as a function $q(\pars)$
of $\pars$:
\begin{eqnarray}
E[\pars] &\equiv& \int \pars\,p(\pars)\,\dd \pars
\\
E[q] &\equiv& \int q(\pars)\,p(\pars)\,\dd \pars
\quad ,
\end{eqnarray}
where again the integrals are implicitly definite integrals over the
entire domain
of $\pars$ (all the parts of $\pars$ space in which $p(\pars)$ is finite) and the
integrals are multi-dimensional if $\pars$ is multi-dimensional.
These expectation values are the mean values of $\pars$ and $q(\pars)$ under
the pdf.  A good sampling---really the definition of a good
sampling---makes the sampling approximation to these integrals
accurate.
With a good sampling $\setof{\pars_k}_{k=1}^K$ the integrals get replaced
with sums over samples $\pars_k$:
\begin{eqnarray}
E[\pars] &\approx& \frac{1}{K}\,\sum_{k=1}^K \pars_k
\\
E[q] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\quad .
\end{eqnarray}
That is, a sampling is good when any expectation value of interest is
accurately computed via the sampling approximation.
The word ``accurately'' here translates into some kinds of theorems
about limiting behavior; the general idea is that the sampling
approximation becomes exact as $K$ goes to infinity.
The size $K$ of the sampling you need \emph{in practice} will depend
on the expectations you want to compute, and the accuracies you need.

As we noted above, in the context of MCMC, we are often sampling or
using some \emph{badly normalized} function $f(\pars)$.
This function is just the pdf $p(\pars)$ multiplied by some unknown and
hard-to-compute scalar.
In this case, for our purposes, the conditions on $f(\pars)$ are that it
be non-negative everywhere and have \emph{finite} integral $Z$
\begin{eqnarray}
0 &\leq& f(\pars) \quad \mbox{for all $\pars$}
\\
Z &=& \int f(\pars)\,\dd \pars
\quad .
\end{eqnarray}

When the sampling $\setof{\pars_k}_{k=1}^K$ is of one of these badly
normalized functions $f(\pars)$---as it usually will be---the
sampling-approximation expectation values are the expectation values
under the properly-normalized corresponding pdf, even though you might
\emph{never learn that normalization}.
When we run MCMC sampling on $f(\pars)$, a function which differs from a
pdf $p(\pars)$ by some unknown normalization constant, then the sampling
permits computation of the following kinds of quantities:
\begin{eqnarray}
E[q] &\equiv& \frac{\int q(\pars)\,f(\pars)\,\dd \pars}{\int f(\pars)\,\dd \pars}
\\
E[q] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\quad .
\end{eqnarray}
That is, the sampling can be constructed (as we will show below) from
$f(\pars)$ directly, and it permits you to compute expectation values
without ever requiring you to integrate either the numerator integral
or the denominator integral, both of which are generally
intractable.\note{As we will
  see, the intractability comes from various directions, but one is
  that the dimension of $\pars$ gets large in most realistic situations.
  Another is that the support of $p(\pars)$ tends to be, in normal
  inference situations, \emph{far smaller} than the full domain of
  $\pars$.
  That is, the pdf is at or very close to zero over all but some tiny
  and very hard-to-find part of the space.}

@@

% in the next paragraph I adopt $D$ to be the dimensionality of the space

For example, one magical property of a sampling (in a $D$-dimensional space)
  is that a \emph{histogram} (a $D$-dimensional histogram) of the samples
  (divided, if you like, by the number of samples in each bin and the bin width\note{something about units and normalization})
  looks very much like the pdf from which the samples were drawn.
This is a way to ``reconstruct'' the pdf from the sampling:
Make a histogram from the samples.
Even in this case, the sampling is being used to do \emph{integrals};
  the (possibly odd) idea is that the approximate or effective value of the pdf in each bin
  of the histogram is an average over the bin.
That average is obtained by performing an integral.

@@

Finally---and perhaps most importantly---a magical property of a sampling
  (in a $D$-dimensional space)
  is that if some of your $D$ dimensions are totally uninteresting
  (nuisance parameters, if you will)
  and some of your $D$ dimensions are of great interest,
  the sampling in the full $D$-space is trivially converted into a sampling in the subspace of interest:
You just drop from each $\pars_k$ vector (or blob or list) the dimensions of no interest!
That is, the projection of the sampling to the subspace of interest produces
  a sampling of the marginalized pdf, marginalizing (or projecting) out the nuisance parameters.%
  \note{This point about marginalization is, once again, a use of the sampling to perform an \emph{integral};
        the marginalized pdf is obtained from the full pdf by an integration.
        The sampling performs this integration automatically.}
That is extremely important for inference,
  where there are always parameters with very different levels of importance to the scientific conclusions.
This point will return again below, either explicitly or implicitly. % HOGG: does it?

@@

In what follows, when we discuss tuning of the proposal pdf, we will
  often cycle through the parameter blob components and propose changes
  in only one dimension or element or component at a time.
That is, at step $k+1$ you might only propose a one-dimensional move
  in the $i$th dimension, not in all $D$ dimensions of the parameter
  space.
That won't change anything significant; but if you are implementing
  this right now you might want to do that.
It will help us tune the proposal pdf (\sectionname~\ref{sec:tuning})
and diagnose problems (\sectionname~\ref{sec:troubleshooting}).

@@

One note to make here
  is that you must \emph{propose}
  in the same coordinates (parameterization or transformation of parameters)
  as that in which your priors are \emph{specified},
  or else multiply in a (possibly nasty) Jacobian.
That is, if your prior is ``flat in $\pars$''
  but you find it easier to run your sampler by proposing steps in $\ln \pars$,
  if you don't modify your acceptance fraction by the Jacobian,
  your \emph{real} prior won't be flat in $\pars$.
These think-os can get subtle;
  the best way to avoid them
  is to write your code and your likelihood function
  and your prior pdf function and your proposal distribution
  all in precisely the same parameterization.
We will return to this point below in our comments on functional testing.

@@

Most MCMC methods
  (though not all\note{We are promoters of an ensemble method that has a
  nearly-fixed proposal distribution and requires no tuning (CITE EMCEE).})
  make use of something like a ``proposal distribution''
  which determines what kinds of steps the walker can take
  as it random-walks through the parameter space.
The user generally has a lot of control over what this proposal distribution might be,
  and how to choose the parameters.
For example, in a $D$-dimensional parameter space, %do we consistently use d here?
  a zero-mean but anisotropic Gaussian (normal distribution)
  is often used to draw offsets for the walker to take from point to point.
Even within this choice (which is of one function among many possibilities),
  there are $D\,(D+1)/2$ parameters in the $D\times D$ symmetric, positive definite covariance matrix
  to set ``by hand''.
How to choose this function and set these parameters?

@@

A very common---%
  and very useful---%
  kind of proposal distribution is one that cycles through parameters,
  taking a random step in just one parameter at a time.
Aside from involving some persistence of state,
  this kind of proposal distribution can be valuable,
  in part because it reduces the (almost impossible) $D$-dimensional tuning problem
  to $D$ one-dimensional problems:
In this form of proposal distrubution,
  it is possible to track a separate acceptance fraction for every parameter.
Code can be built that uses the burn-in phase to tune all $D$ proposal variances
  such that each of them, individually,
  obtains acceptance at the same Goldilocks ratio.

@@

Another proxy for autocorrelation time useful for tuning is the
  Expected Squared Jump Distance.\note{(CITE GELMAN)}
This is the mean squared distance the walker moves, per step.
It is maximized when the acceptance ratio is reasonable
  and the step size is large;
  it is large when the exploration of the space is fast.
This proxy is easy to measure and use for tuning,
  and more directly related to autocorrelation time than the acceptance ratio.
We recommend using it for tuning, though we have never used it ourselves.
Like the acceptance fraction, it can also be used to tune
  in the case that the proposal distribution loops over parameters;
  once again, the user gets (in this case) $D$ one-dimensional tunings.

@@

As we mentioned above, scientists often like to report the ``best sample'';
  that is, the sample with the highest posterior pdf value.
This is not usually a good idea.
For one, MCMC is a sampling algorithm, not an optimization algorithm;
  there are no guarantees that it will find the optimum of the posterior pdf in reasonable time.%
  \note{The sampler samples the pdf fairly.
    The tallest peak in the posterior is not guaranteed to be---%
    and in general is not---%
    also the peak with the greatest posterior ``mass''.
    That is, the samples in a fair sampling will not necessarily come
    predominantly from the \emph{tallest} peak in the posterior pdf.
    They will come from the peak with the greatest total integral
    of the likelihood over the prior.
    These issues might sound ``academic'',
    but when the number of parameters gets large (greater than, say, 10),
    many normal intuitions one might have (if any) about what is reasonably likely in a pdf
    are regularly violated.}
For two, the closeness of the best sample to the posterior mode is a strong function of sample size,
  and with very bad scaling.%
  \note{If you have a $K$-point sampling in a $D$-dimensional parameter space,
    there is some best sample $\pars_k^{\best}$.  Now imagine that you want to make
    that best sample 10 times better---that is, you want to find a point $\pars^{\better}$
    that is ten times closer to the true optimum.  How many more samples do you need to take?
    In the limit that $K$ is large, on average, you have to take $10^D$ more samples!
    Most sensible optimizers are far, far better than this.}
For three, if you just want an optimum, use an optimizer!%
  \note{HOGG, something?}
That is, we \emph{strongly advise against} reporting,
  as ``the'' measurement or ``the result'',
  the parameter value or values for the best sample.
Giving some posterior samples is a good idea (see below).
If you feel drawn to give the best sample,
  instead give the parameter values found by optimizing the posterior PDF,
  starting at the best sample as an initialization
  (and sacrifice your probabilism\note{something}).
This optimal-posterior parameter value is called the
  ``maximum \foreign{a posteriori}'' or ``MAP'' value for the parameters;
  it is like a maximum-likelihood value but regularized by the prior.
It is an estimator with some good (and some bad) properties, but it is not the point of MCMC,
  and therefore outside the scope of this \documentname.

@@

MCMC is used to obtain samples $\pars_k$ from a pdf $p(\pars)$,
  or a badly normalized all-positive function $f(\pars)$
  that is different from the pdf by an unknown factor $Z$.
In the context of data analysis,
  MCMC is usually being used to obtain samples $\pars_k$
  that are \emph{parameter values} $\pars$ for a probabilistic model of some data.
That is, MCMC is sampling a pdf \emph{for the parameters}.
If it is returning parameter values, then it is not sampling a pdf for the data.
That is, MCMC can only sample a posterior (or prior) pdf for the parameters.
\emph{It cannot sample a likelihood.}%
\note{Well, technically, MCMC \emph{can} be used to sample a likelihood function,
  but the samples would be samples of possible \emph{data} not possible \emph{parameters}.
  The purist might say that even this is not sampling a likelihood function,
  because you should only call it a ``likelihood function'' in contexts in which you are
  treating the data as fixed (at the true data values) and the parameters as variable.
  In this context, the likelihood function is \emph{not} a pdf for the parameters,
  so it can't be sampled.}

@@

% Hogg say:  I have no idea what the comment here (below) means, even though I wrote it!
% ---
% Give a sense of the difficulties you get into if you fail here; like
% discussion with Sven Kreiss on 2013-02-14...
% ---

Ideally, if you are using MCMC for inference, your code should input
to the MCMC function or routine a probability function which is a
product of a prior times a likelihood.
In terms of implementation, it is usually advisable to work in the
logarithm of the density function, so the function input to the MCMC code
would be called something like \code{ln_f()}.
This function \code{ln_f()} internally would compute and return the
sum of a log prior pdf \code{ln_prior()} and a log likelihood function
\code{ln_likelihood()}.
(When working in logarithmic probability, in M--H-like MCMC methods,
the accept--reject step would change from a comparison of a random
deviate with a ratio of probabilities to comparison of the log of a
random deviate with a difference of log probabilities.)

If you are using ``flat'' (improper) priors, the \code{ln_prior()}
function can just return a zero no matter what the parameters.
If it is flat with bounds (that is, proper), the \code{ln_prior()}
should check the bounds and return \code{-Inf} values if the parameter
vector is out of bounds.
The pseudo-code for the \code{ln_f()} function should look something
like this:
\begin{verbatim}
def ln_f(pars, data):
    x = ln_prior(pars)
    if is_Inf(x):
        return -Inf
    return x + ln_likelihood(data, pars)
\end{verbatim}
This pseudo-code ensures that you only compute the likelihood when the
parameters are within the prior bounds; it assumes implicitly that the
prior pdf is easier to compute than the likelihood.
If you find yourself in the opposite regime, adjust accordingly.

@@

\paragraph{functional testing:}
In addition to a full set of unit tests for every part of your code
  (yeah \emph{right}\note{If you aren't unit testing, you are probably making some very big mistakes.}),
  we recommend building some end-to-end functional tests that permit you to perform MCMC
  runs with output that must meet certain expectations.
The simplest kind of functional test is to sample a known distribution,
  and check that the sampling has the moments you expect (to within tolerances).
For example, you can sample a $D$-dimensional Gaussian distribution with known non-trivial covariance,
  and check that the empirical covariance comes out as expected.
This is an easy test to fail, so a success builds confidence in any MCMC code.

In the most extreme form of this test,
  for a Gaussian in $D$ dimensions,
  build the covariance matrix input to the Gaussian distribution
  by taking the sum of the outer product of $D+1$ randomly generated vectors.
Then, before sampling, get the eigenvalues and eigenvectors of the matrix.
Set up the MCMC sampler with the Gaussian function with this covariance matrix
  as the probability function (or the log of it as the log-probability function).
Sample for many autocorrelation times,
  or until you are confident you are getting good samples.
Check that the empirical $D$-dimensional covariance of the samples
  has the same (or similar) eigenvectors and eigenvlues to the input matrix.
Also check that the individual one-dimensional samplings
  (along each axis or any projection in the $D$-space)
  look Gaussian and have nearly vanishing skew and kurtosis.
It takes some thinking and some calculating to see ``how well'' these things all ought to match
  expectations, but one thing that can be said with confidence is that agreement should improve
  as the sampling proceeds.
If it doesn't, it is likely that there is something broken about the sampler itself.
